Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-05-19 manuscript/manuscript.Rmd
updates ms with error results

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2012]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes such as survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may be difficult to resolve in statistical models based on population-level data where demographic processes are not identifiable. If important climate effects are missed because of the aggregation inherent in in population-level data, then population models built with such data will make uninformative or unreliable forecasts.

Here, we compare the forecasting skill of statistical and population models based on aggregated, population-level data with models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. We use the models to answer two questions: (1) Can population models fit using aggregated individual-level data (percent cover) produce forecasts as skillful as those from models fit to demographic data? 
And (2) Can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data?

@@

Results
-------
Both the IPM and QBM generated skillful one-step-ahead forecasts for out-of-sample observations, with an average correlation between predictions and observations ($\rho$) of 0.# across all models (Fig. 2).
Without climate covariates, the accuracy of forecasts from the IPM were not statistically greater than the accuracy of forecasts from the QBM.
With climate covariates, the best out-of-sample predictive model (highest *lppd*) for each species and vital rate typically resulted from highly constrained priors on the climate effects (Fig. S1).
Thus, the posterior distributions of climate effects included in our models overlapped zero and generally shrunk toward zero, though for some species-vital rate combinations important effects (80% credible interval does not include zero) did emerge (Fig. 3). 
Despite the small effects, including climate covariates did increase the accuracy of forecasts for all species except *P. smithii* (Fig. 2).
However, increases in accuracy due to the inclusion of climate covariates were not significant (*P* > 0.05 for all comparisons of $\rho$ between climate and no-climate forecasts within model types; Fig. 2).
In only one case were IPM forecasts significantly more accurate than the QBM (Fig. 2): forecast accuracy of *P. secunda* percent cover from an IPM with climate covariates was greater than the accuracy from an equivalent QBM ($t_{(195)}$ = 1.72, *P* = 0.043).
The accuracy of both model's forecasts declined as the forecast horizon increased, but they did so at similar rates (Fig. 4).
The only exception is for *P. secunda* where forecast accuracy appears to slightly increase with forecast horizon, after an initial decrease from a forecast horizon of one year (Fig. 4).

@@

Discussion
----------
Population models built using individual-level data allow inference on demographic processes, but they can only forecast future population states across the (typically limited) spatial extent of the observations. 
Population-level data are much easier to collect across broad spatial extents, so models built using such data offer an appealing alternative to traditional population models [@Queenborough2011]. 
However, density-structured models rely on the aggregation of individual-level data. 
This creates a potential problem if (1) such models inaccurately estimate density-dependence or (2) are to be used in a climate change context because it is individuals, not populations, that respond to climate [@Clark2012]. Can models based on population-level metrics generate forecasts that are as skillful as those generated from models based on individual-level data? Are models based on population-level metrics as sensitive to climate as models based on individual-level data? Do we need detailed demographic data to forecast the state of plant populations?

@@

We also expected IPM forecast accuracy to decline at a lower rate than the QBM as the forecast horizon increased.
Indeed, many have argued that more mechanistic models produce better predictions, especially under novel conditions [@Evans2012b; @Schindler2015b].
Likewise, the IPM explicitly models the processes of recruitment and survival that may be poorly represented in the QBM since recruitment and survival mainly affect small plants that contribute little to overall changes in percent cover.
But the addition and subtraction of small plants can have large effects on population growth over time and potentially lead to a longer forecast horizon for models that better represent the plant life cycle.
We found no evidence for a difference between the IPM and QBM forecast horizons (Fig. 4).
Similar forecast horizons should be expected if both models adequately capture density-dependence and environmental forcing is negligble, as is the case for our models (Fig. 2).
In fact, only the forecast horizons for *P. secunda* noticeably differ (Fig. 4), and that is also the only species where we found the IPM to outperform the QBM when we included climate effects.

>>>>

Addressing climate change questions with models fit to population-level data is potentially problematic. It is individuals, not populations, that respond to climate variables [@Clark2011b]. Ignoring this fact amounts to an "ecological fallacy", where inference about the individual relies on statistical inference on the group [@PIANTADOSI1988]. Population growth (or decline) is the outcome of demographic processes such as survival, growth, and recruitment that occur at the level of individual plants. Climate can affect each demographic process in unique, potentially opossing, ways [@Dalgleish2011]. These unique climate responses may be difficult to resolve in statistical models based on population-level data where demographic processes are not identifiable. Futhermore, models based on aggregated data may reflect short-term effects of one vital rate (e.g., plant growth) more than others, but over the long-term the influence of other vital rates may be important. If important climate effects are missed because of the aggregation inherent in in population-level data, then population models built with such data will make uninformative or unreliable forecasts.

Here, we compare the forecasting skill of statistical and population models based on aggregated, population-level data with models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. We use the models to answer two questions: (1) Can population models fit using aggregated individual-level data (percent cover) adequately capture density dependence to produce forecasts as skillful as those from models fit to demographic data? 
And (2) Can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data? 
Both questions are motivated by the fact that the effects of intraspecific competition (density dependence) and interannual weather variability act at the level of the individual [@Clark2011b], meaning that models based on individual-level processes should generate more skillful forecasts.

@@

Results
-------
Both the IPM and QBM generated skillful one-step-ahead forecasts for out-of-sample observations, with an average correlation between predictions and observations ($\rho$) of 0.73 across all models and species (Fig. 2).
Without climate covariates, the accuracy of forecasts from the IPM were not statistically greater than the accuracy of forecasts from the QBM (Fig. 2) and overall error was similar (mean absolute error; Fig. S1).
With climate covariates, the best out-of-sample predictive model (highest *lppd*) for each species and vital rate typically resulted from highly constrained priors on the climate effects (Fig. S2).
Thus, the posterior distributions of climate effects included in our models overlapped zero and generally shrunk toward zero, though for some species-vital rate combinations important effects (80% credible interval does not include zero) did emerge (Fig. 3). 
Despite the small effects, including climate covariates did increase the accuracy of forecasts for all species except *P. smithii* (Fig. 2).
Increases in accuracy due to the inclusion of climate covariates were not significant (*P* > 0.05 for all comparisons of $\rho$ between climate and no-climate forecasts within model types; Fig. 2).
However, the IPM with climate covariates for *P. secunda* did have significantly lower error than the IPM without climate covariates ($t_{(196)}$ = -1.84, *P* = 0.033; Fig. S2).
In only one case were IPM forecasts significantly more accurate than the QBM (Fig. 2): forecast accuracy of *P. secunda* percent cover from an IPM with climate covariates was greater than the accuracy from an equivalent QBM ($t_{(195)}$ = 1.72, *P* = 0.043).
Forecasts from IPMs with climate covariates had significantly lower error than equivalent QBM forecasts for *P. smithii* ($t_{(216)}$ = -3.49, *P* < 0.001) and *P. secunda* ($t_{(196)}$ = -1.83, *P* = 0.034) (Fig. S2).
Results from all pairwise statistical tests are shown in Table S1.
The accuracy of both model's forecasts declined as the forecast horizon increased, but they did so at similar rates (Fig. 4).
The only exception is for *P. secunda* where forecast accuracy appears to slightly increase with forecast horizon, after an initial decrease from a forecast horizon of one year (Fig. 4).

@@

Discussion
----------
Population models built using individual-level data allow inference on demographic processes, but they can only forecast future population states across the (typically limited) spatial extent of the observations. 
Population-level data are much easier to collect across broad spatial extents, so models built using such data offer an appealing alternative to traditional population models [@Queenborough2011]. 
However, density-structured models rely on the aggregation of individual-level data. 
This creates a potential problem if (1) such models inaccurately estimate density dependence or (2) are to be used in a climate change context because it is individuals, not populations, that respond to intraspecific competition and weather [@Clark2011b]. Can models based on population-level metrics generate forecasts that are as skillful as those generated from models based on individual-level data? Are models based on population-level metrics as sensitive to climate as models based on individual-level data? Do we need detailed demographic data to forecast the state of plant populations?

@@

We also expected IPM forecast accuracy to decline at a lower rate than the QBM as the forecast horizon increased.
Indeed, many have argued that more mechanistic models produce better predictions, especially under novel conditions [@Evans2012b; @Schindler2015b].
Likewise, the IPM explicitly models the processes of recruitment and survival that may be poorly represented in the QBM since recruitment and survival mainly affect small plants that contribute little to overall changes in percent cover.
But the addition and subtraction of small plants can have large effects on population growth over time and potentially lead to a longer forecast horizon for models that better represent the plant life cycle.
We found no evidence for a difference between the IPM and QBM forecast horizons (Fig. 4).
Similar forecast horizons should be expected if both models adequately capture density dependence and environmental forcing is negligble, as is the case for our models (Fig. 2).
In fact, only the forecast horizons for *P. secunda* noticeably differ (Fig. 4), and that is also the only species where we found the IPM to outperform the QBM when we included climate effects.
